---
title: "Getting started with halosimr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with halosimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 3.5, dpi = 96
)
```

`halosimr` simulates exposure to high-acuity, low-occurrence (HALO) events
across a population of providers, models how readiness decays between
exposures, and estimates how training programs change it. It extends the
methodology of the Code Blue Blindspots paper (PMID 41633464).

## The idea

Some clinical events are rare enough that a given provider may see only a few in
a year, yet critical when they occur, and skill at handling them fades with time
since the last one. `halosimr` asks how a provider population's preparedness for
these HALO events rises and falls over a year, and how scheduling and training
change it.

The model runs on one chain. Events occur on random days. A provider is
**exposed** when they are on shift the day an event happens. Each exposure
**resets** that provider's **readiness** (a 0-to-1 proxy for skill, 1 right after
an exposure) to full; between exposures readiness **decays**. The run of days
between a provider's consecutive exposures is a **gap**, and a long gap means low
readiness when the next event arrives. A **training** session acts as a scheduled
reset, topping readiness back up without waiting for a real event.

So a study has three ingredients:

1. a **schedule** of who works which shifts,
2. a stream of **HALO events**,
3. a **readiness model** that sets how fast skill decays between resets.

```{r setup}
library(halosimr)
```

## Build the inputs

Generate a schedule and a stream of events. `schedule_type = "3/7 Day"` gives
each provider three day-shifts per seven-day week; `rate = 0.1` means about 0.1
HALO events per day, roughly 36 across the year.

```{r inputs}
sched <- generate_schedule(
  n_providers = 200, n_days = 365,
  schedule_type = "3/7 Day", seed = 42
)
ev <- generate_events(n_days = 365, rate = 0.1, seed = 42)

dim(sched)         # providers x days; cells "d" (Day), "n" (Night), "o" (Off)
head(ev)           # one row per event
```

The `"Random"` schedule type draws each day from the empirical Day, Night, and
Off weights reported in the Code Blue Blindspots paper (Day = 0.246,
Night = 0.230, Off = 0.524). Pass `weights` to override, for example
`weights = c(d = 1/3, n = 1/3, o = 1/3)` for an equal split.

## Run one simulation

`halo_sim()` bundles the inputs and configuration; `hsrun()` executes the
pipeline and attaches the results. Each readiness model reads its own
parameters: `exponential` uses `readiness_half_life_days` (days for readiness to
halve). `readiness_threshold_days` is separate from the decay curve; it sets the
gap length considered too long, used for the flag below.

```{r run}
sim <- halo_sim(
  sched, ev,
  readiness_model = "exponential",
  readiness_half_life_days = 60,
  readiness_threshold_days = 90
)
sim <- hsrun(sim)
sim
```

The summary gives the average readiness among on-shift providers and the share
of providers who had at least one gap longer than `readiness_threshold_days`, a
flag for who went dangerously long without an exposure.

`results_df` summarises each provider's gaps (the day-counts between their
consecutive exposures). Key columns are `n_events` (exposures in the window),
`gap_max` (their longest gap), and `max_gap_exceeds_threshold` (whether that
longest gap ran past the threshold):

```{r results}
head(sim$results_df)
```

`proportion_ready_on_shift`, the daily mean readiness across providers working
that day, is the primary outcome.

## Single-run figures

`plot_readiness` traces the on-shift readiness curve across the year;
`plot_gap_distribution` shows how provider maximum gaps are distributed, with the
threshold marked.

```{r single-plots}
plot_readiness(sim)
plot_gap_distribution(sim)
```

## Compare training programs with Monte Carlo

A single run reflects one random draw of events, so its curve is noisy.
`hsrunmc()` repeats the run over `n` different random seeds and summarises the
day-by-day spread: the median readiness across runs, plus a shaded band for the
10th-to-90th-percentile range. Each seed needs its own fresh schedule and
events, so instead of fixed inputs you pass two functions that take a seed and
build the inputs for that run. When a training program is set, `hsrunmc()` also
runs an untrained baseline arm and reports the lift in readiness: the per-day
gain of the trained arm over the baseline.

```{r mc}
mc <- hsrunmc(
  n = 30,
  schedule_fn = function(s) generate_schedule(200, 365, "3/7 Day", seed = s),
  events_fn   = function(s) generate_events(365, rate = 0.1, seed = s),
  readiness_model = "exponential", readiness_half_life_days = 60,
  training_program = "monthly"
)
mc
```

```{r mc-plot}
plot_mc_band(mc)
```

The trained band sits above the baseline; `mc$lift` is the per-day difference in
median readiness.

## Readiness models

Four decay kernels are available and can be called directly on a vector of days
since the last reset:

```{r kernels}
days <- c(0, 30, 60, 90, 180)
rbind(
  binary      = readiness_binary(days, threshold = 90),
  exponential = readiness_exponential(days, half_life = 60),
  ebbinghaus  = readiness_ebbinghaus(days, b = 0.05),
  step        = readiness_step(days, t1 = 90, t2 = 180, partial = 0.5)
)
```

Select one in a simulation with `readiness_model` plus its parameters, for
example `halo_sim(..., readiness_model = "step", step_t2_days = 180)`.

## Reports

For a shareable document, render the bundled template, which runs a full
single-run plus Monte Carlo analysis from parameters you supply. Pass
`output_file` to control where the HTML lands; without it the output is written
next to the template inside the installed package.

```{r report, eval = FALSE}
rmarkdown::render(
  system.file("report_template.Rmd", package = "halosimr"),
  output_file = path.expand("~/Desktop/halo_report.html"),
  params = list(n_providers = 200, n_days = 365, training_program = "monthly")
)
```
